---
title: "bm_final_project"
author: "Yun He"
date: "December 6, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 

```{r}
cancer = read_csv("./Cancer_Registry.csv") %>% 
 mutate(medIncome = as.numeric(medIncome)) %>% 
  mutate(income_cutoff = ifelse(medIncome >= 47000,1,0)) %>% 
  separate(Geography, into = c("County", "State"), sep = ",") %>% 
  select(-PctSomeCol18_24) # ‘PctSomeCol18_24’ column should not be included because there are too many missing values.
  colSums(is.na(cancer)) # there is no missing values 
cancer = cancer %>% 
  select(-binnedInc, -PctNoHS18_24, -avgAnnCount, -avgDeathsPerYear) %>% 
  group_by(State) 
```
'avgAnnCount' and 'avgDeathsPerYear' should be removed because we are not measuring the average value. 'binnedInc' should be removed because we have already set the cutoff for median income. ‘PctSomeCol18_24’ column should not be included because there are too many missing values. 'PctNoHS18_24' should be removed because there is no corresponding variable for people who are 25 and over.
