---
title: "bm_final_project"
author: "Yun He"
date: "December 6, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(glmnet)
library(HH)

library(faraway)
library(leaps)
library(caret)
library(boot)
library(broom)

```

```{r}
cancer = read_csv("./Cancer_Registry.csv")
colSums(is.na(cancer)) # 'PctSomeCol18_24', 'PctEmployed16_Over','PctPrivateCoverageAlone' should be removed because they have lots of NA.  
```

```{r}
cancer = read_csv("./Cancer_Registry.csv") %>% 
  select(-PctSomeCol18_24, -PctEmployed16_Over, -PctPrivateCoverageAlone) %>% 
  # set the cutoff for low and high income as 47k
  mutate(medIncome = as.numeric(medIncome)) %>% 
  mutate(income_cutoff = ifelse(medIncome >= mean(cancer$medIncome),1,0)) %>%
  # set the cutoff for age at mean median age 
  mutate(age_cutoff = ifelse(MedianAge >= mean(cancer$MedianAge),1,0)) %>%
  separate(Geography, into = c("County", "State"), sep = ",") %>%
  # remove unnecessary columns 'binnedInc'sicne we have already set the cutoff for income and 'PctNoHS18_24' should be removed because there is no corresponding variable for people who are 25 and over.
  select(-binnedInc, -PctNoHS18_24) %>% 
  # make dependent variable at 1st column 
  select(TARGET_deathRate, everything())
  # see the correlation
cancer_cor = cancer %>% 
  select(-State, -County)
cor(cancer_cor) %>% knitr::kable()
  # get descriptive information of cancer 
attach(cancer)
summary(cancer)
```

```{r}
 # make multilinear regression model to see which variables are significant 
multi.fit <- lm(TARGET_deathRate ~ ., data = cancer_cor)
summary(multi.fit)
```

```{r}
  # select 10 variables that we choosen after compare with the information from the research literature 
cancer_model = cancer %>% 
  select(TARGET_deathRate, avgAnnCount, incidenceRate, AvgHouseholdSize, studyPerCap, income_cutoff, age_cutoff, PercentMarried, PctPrivateCoverage, PctWhite, PctHS25_Over)  
 # make histogram to see the distribution of the variables and decide whether we need transformation 
cancer_model %>%
  select(-income_cutoff, -age_cutoff) %>%
  gather(measure, value) %>%
  ggplot(aes(value)) +
  facet_wrap(. ~ measure, scales = "free") +
  geom_histogram()
 # all the correlation value are below 0.7, so there is no covariances. 
 cor(cancer_model) %>% knitr::kable()
```

## Data cleaning 

```{r}
cancer = read_csv("./Cancer_Registry.csv") %>% 
  janitor::clean_names() %>% 
  separate(geography, into = c("county", "state"), sep = ",")

## Check with missing values
skimr::skim(cancer %>% select(-state, -county))

## Select variables
cancer = cancer %>% 
  dplyr::select(-avg_ann_count, -avg_deaths_per_year, -pop_est2015, -med_income, -poverty_percent, -median_age_male, -median_age_female, -avg_household_size, -pct_some_col18_24, -pct_private_coverage_alone, -birth_rate, -pct_married_households) %>% 
  dplyr::select(state, county, target_death_rate, everything()) %>% 
  mutate(binned_inc = factor(binned_inc),
         binned_inc = relevel(binned_inc, ref = "[22640, 34218.1]"),
         ct_status = ifelse(study_per_cap == 0, 0, 1),
         age_cat = ifelse(median_age > 20 & median_age <= 39, 0, 1)) %>% 
  dplyr::select(-study_per_cap, -median_age)
```

The original dataset has 34 variables and 3047 observations. After exploring the dataset, we found that variables 'pct_employed16_over', 'pct_private_coverage_alone' and 'pct_some_col18_24' had missing values. Variables 'pct_private_coverage_alone' and 'pct_some_col18_24' has to be removed for subsequent analyses because of the relatively large amount of missing values. We chose the variable 'binned_inc' from 'med_income', 'poverty_percent' and 'binned_inc' for measuring the income level. We chose 'percent_married' from 'percent_married' and 'pct_married_households' for measuring the married status. We also chose 'incidence_rate' from 'avg_ann_count' and 'incidence_rate' for measuring mean cancer diagnoses annually. We chose 'median_age' from 'median_age', 'median_age_female' and 'median_age_male' for measuring the age. In addition, because our outcome is 'target_death_rate' which can be calculated from dividing the variable 'avg_deaths_per_year' by the variable 'pop_est2015', we removed these two variables from our dataset. 'avg_household_size' and 'birth_rate' seem to be uncorrelated with the outcome, thus, we removed them from our dataset as well. We ended with the dataset of 23 variables and 3047 observations.

## Data exploration

```{r}
## Distribution of continuous variables
descp_stats = function(x){
  df = broom::tidy(summary(x)) %>% 
    mutate(sd = sd(x))
  return(df)
}

cancer %>% 
  dplyr::select(-state, -county, -binned_inc, -ct_status, -age_cat) %>% 
  map_df(descp_stats) %>%
  mutate(variable = c("target_death_rate", "incidence_rate", "percent_married", "pct_no_hs18_24", "pct_hs18_24", "pct_bach_deg18_24", "pct_hs25_over", "pct_bach_deg25_over", "pct_employed16_over", "pct_unemployed16_over", "pct_private_coverage", "pct_emp_priv_coverage", "pct_public_coverage", "pct_public_coverage_alone", "pct_white", "pct_black", "pct_asian", "pct_other_race")) %>% 
  dplyr::select(variable, mean, sd, everything()) %>%
  knitr::kable(digits = 2)

## Distribution of the death rate
cancer %>%
  ggplot(aes(x = target_death_rate)) + 
  geom_histogram(aes(y = ..density..)) + geom_density(color = "red")

## Distribution of categorical variables
cancer %>% 
  ggplot(aes(x = binned_inc, y = target_death_rate)) + geom_boxplot() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

cancer %>% 
  ggplot(aes(x = factor(ct_status), y = target_death_rate)) + geom_boxplot()

cancer %>% 
  ggplot(aes(x = factor(age_cat), y = target_death_rate)) + geom_boxplot()
```

The outcome 'death rate' is normally distributed.

## Fit the model

```{r}
cancer_mod = cancer %>% 
  dplyr::select(-state, -county)

fit = lm(target_death_rate ~ ., data = cancer_mod)
summary(fit)

step(fit, direction = "backward")
refit = lm(target_death_rate ~ incidence_rate + binned_inc + 
    pct_no_hs18_24 + pct_hs18_24 + pct_hs25_over + pct_bach_deg25_over + 
    pct_employed16_over + pct_private_coverage + pct_emp_priv_coverage + 
    pct_public_coverage + pct_public_coverage_alone + pct_white + 
    pct_other_race, data = cancer_mod)
vif(refit)

model_fit = lm(target_death_rate ~ incidence_rate + binned_inc + 
    pct_no_hs18_24 + pct_hs18_24 + pct_hs25_over + pct_bach_deg25_over + 
    pct_employed16_over + pct_emp_priv_coverage + pct_white + 
    pct_other_race, data = cancer_mod) 
summary(model_fit)
plot(model_fit)



```

## Lasso regression

```{r}
cancer = cancer %>% 
  dplyr::select(-avg_ann_count, -avg_deaths_per_year, -pop_est2015, -binned_inc, -poverty_percent, -median_age_male, -median_age_female, -avg_household_size, -pct_some_col18_24, -pct_private_coverage_alone, -birth_rate, -pct_married_households) %>% 
  dplyr::select(state, county, target_death_rate, everything()) %>% 
  mutate(ct_status = ifelse(study_per_cap == 0, 0, 1),
         age_cat = ifelse(median_age > 20 & median_age <= 39, 0, 1)) %>% 
  dplyr::select(-study_per_cap, -median_age)

cancer_mod = cancer %>% 
  dplyr::select(-state, -county, -pct_employed16_over) %>% 
  dplyr::select(target_death_rate, incidence_rate, med_income, pct_no_hs18_24, pct_hs18_24, pct_hs25_over, pct_bach_deg25_over, pct_emp_priv_coverage, pct_white, pct_other_race)

## no NA
Y <- as.matrix(cancer_mod[,1])
X <- as.matrix(cancer_mod[,-1])
grid <- 10^seq(5,-2, length=100)
ridge3<-glmnet(X, Y, alpha=1, lambda=grid)

set.seed(2)
cv.out<-cv.glmnet(X,Y)
plot(cv.out)
best.lambda<-cv.out$lambda.min
lasso2<-glmnet(X, Y, alpha =1, lambda=best.lambda)
coef(lasso2)
```

